### 2015-01-20 PLDI15_Response/response.txt

We would like to thank the reviewers for their insightful comments. We first
address the concerns raised my multiple reviewers first followed by individual
concerns.

Complexity of contracts
-----------------------
We tend to disagree with the idea that the difficulty of expressing the
contracts is the same as choosing the correct store consistency level. While
the programmer does need to translate the integrity constraints in the
application to consistency constraints, the advantage is that the programmers
write contracts for the abstract system model. This results in generic RDTs
that can be realized on different data stores (as our implementation
illustrates) with varied consistency lattices.

The Quelea contract language itself is built upon three primitive,
well-understood relations -- vis, so and sameobj, along with the usual logical
and relational operators. Moreover, in the presence of coordination-free
transactions with weaker isolation levels, the choice of correct consistency
level is non-trivial. Quelea incorporates transactions with a single additional
primitive relation sametxn. We believe the apparent complexity of contracts can
be further reduced using syntactic sugar, similar to our relational operators.

Comparison to previous work
---------------------------
Our work varies significantly from the previous works Bailis2013b and
Burckhardt2014. While Quelea provides a generic framework to describe new RDTs
and verifiably compile them to existing data stores, Bailis2013b is concerned
with implementing stronger consistency levels on top of existing stores.
Bailis2013b is not concerned with language abstractions, but rather an
implementation strategy. Our implementation, while sharing similarities with
Bailis2013b, illustrates that the Quelea programming model can be efficiently
realized on existing data stores. Moreover, Bailis2013b did not incorporate
transactions, which Quelea does.

Burckhardt2014 describes a formal framework for expressing, verifying and
comparing different forms of eventual consistency, and is not meant for
practical programming over ECDS like Quelea. Moreover, unlike Burckhardt2014,
Quelea focuses on automatic verification of contracts using SMT solver, which
is essential for practicality. Quelea additionally incorporates support for
programming with coordination-free transactions.

Compile Time
------------
Our contract classification through the solver took on average took 11.5 ms
across the benchmarks. We expect this SMT solver overhead to remain low as the
application scales. This is because we do not analyze the RDT/transaction
implementation to obtain the verification conditions but rather simply analyze
the user provided contracts. The contracts are encoded as SMT formulas. In
particular, each of our four primitive relations is encoded as an uninterpreted
function. Hence, the state space for the validity check remains low. Moreover,
our language is defined carefully such that the validity check is decidable for
which efficient decision procedures exist.

The time to classify a contract is proportional to the number of consistency
levels in the partially ordered lattice. For example, Doug Terry et al.'s
Session Guarantees for Weakly Consistent Replicated Data introduced 4
incomparable session guarantees, and the consistency level for an operation is
any combination of the 4 session guarantees. The classification here, at worst,
requires 4! = 24 comparisons (SMT validity tests). Since each comparison is
very cheap, the solution remains scalable.

Review A
--------
The challenge with a dynamic strategy for enforcing contracts is the
translation of contracts, which axiomatically describe correct execution to a
validity check over the dynamic execution state. Such a translation is
non-obvious.

We used the same number of shim layer nodes for the results in Figure 9 a,b,c
-- 5. Shim Layer nodes are soft-state entities and can be elastically scaled on
demand. All the benchmarks except Twissandra and RUBiS were written from
scratch.

Review B
--------
Bounded staleness could be accommodated into the contract language, but the
contract classification becomes undecidable since the theory of integer
arithmetic is undecidable.

RDT specification language does to some extent guards against
malicious/incorrect code. In particular, RDT operations are side-effect free.
While it is possible that the reduction can end up not terminating, it cannot
do any more harm.

A crashed operation does not affect the durable log. An operation is added to
the log only if it is successful. An operation is considered successful if it
produces a result.

R+ could be a total relation, but the R+ is still useful because it is not
required to be a total relation. For example, consider causal
consistency (CC) contract which is defined as:

 hbo \subseteq vis

where hbo = (soo U vis)+. Because of our lenient encoding of “+", hbo can be
any superset of  transitive closure of (soo U vis). Nevertheless, all we can
assume about hbo is that it contains the  transitive closure of (soo U vis);
there is no proof of membership for other elements in hbo.  Consequently, by
asserting hbo \subseteq vis, the strongest guarantee that causal consistency 
contract is offering is that the transitive closure of (soo U vis) is contained
in vis. An application  contract (\psi) which requires more than the transitive
closure of (soo U vis) to be contained in vis  cannot be satisfied under CC
because the following implication is simply not valid even under our lenient
definition of “+“ :

 \Delta |- (hbo \subseteq vis) => \psi

\Delta, as usual, denotes the conjunction of basic axioms about primitive
relations.

Review E
--------
Our idea is to parameterize the classification procedure for each target data
store. Hence, if multiple partially ordered levels are appropriate, the
classification procedure can be encoded to choose the desired consistency
level based on external knowledge such as price.

### 2015-01-20 PLDI15_Response/response.txt

We would like to thank the reviewers for their insightful comments. We first
address the concerns raised my multiple reviewers first followed by individual
concerns.

Complexity of contracts
-----------------------
We tend to disagree with the idea that the difficulty of expressing the
contracts is the same as choosing the correct store consistency level. While
the programmer does need to translate the integrity constraints in the
application to consistency constraints, the advantage is that the programmers
write contracts for the abstract system model. This results in generic RDTs
that can be realized on different data stores (as our implementation
illustrates) with varied consistency lattices.

The Quelea contract language itself is built upon three primitive,
well-understood relations -- vis, so and sameobj, along with the usual logical
and relational operators. Moreover, in the presence of coordination-free
transactions with weaker isolation levels, the choice of correct consistency
level is non-trivial. Quelea incorporates transactions with a single additional
primitive relation (sametxn). We believe the apparent complexity of contracts can
be further reduced using syntactic sugar, similar to our relational operators.

Comparison to previous work
---------------------------
Our work varies significantly from the previous works Bailis2013b and
Burckhardt2014. <!-- While Quelea provides a generic framework to describe new RDTs
and verifiably compile them to existing data stores, Bailis2013b is concerned
with implementing stronger consistency levels on top of existing stores.
Bailis2013b is not concerned with language abstractions, but rather an
implementation strategy. Our implementation, while sharing similarities with
Bailis2013b, illustrates that the Quelea programming model can be efficiently
realized on existing data stores. Moreover, Bailis2013b did not incorporate
transactions, which Quelea does. -->
<!-- GK: 1. I think the statement "a generic framework to describe new RDTs and
verifiably compile them to existing data stores" does not sufficiently
distinguish Quelea from Bailis2013b. -->

Bailis2013b describes implement strategies to realize stronger consistency
levels on top of eventually consistent stores. In contrast, Quelea provides a
language framework to describe application-level consistency requirements and
automatically map them to *some* implementation of appropriate store consistency
level. As such, Quelea's contract classification is independent of how various
consistency levels are actually implemented in the data store.  Furthermore,
Bailis2013b did not incorporate transactions, which Quelea does. Nevertheless,
our own implementation of strong consistency levels on top of Cassandra does
bear some similarities with Bailis2013b.

@@

Burckhardt2014 describes a formal framework for expressing, verifying and
comparing different forms of eventual consistency, and is not meant for
practical programming over ECDS like Quelea. Moreover, unlike Burckhardt2014,
Quelea focuses on automatic verification of contracts using SMT solver, which
is essential for practicality. Quelea additionally incorporates support for
programming with coordination-free transactions.

Compile Time
------------
Our contract classification through the solver took on average took 11.5 ms
across the benchmarks (this statistic is also available on p.11 of the paper).
We expect this SMT solver overhead to remain low as the application scales. This
is because we do not analyze the RDT/transaction implementation to obtain the
verification conditions but rather simply analyze the user provided contracts.
Further, our contract language is defined carefully such that the validity check
is decidable for which efficient decision procedures exist. Moreover, our
contract language artifacts (e.g: primitive relations, effects), are encoded as
uninterpreted symbols in SMT without the need for expensive theories, such as
linear arithmetic.  Hence, the state space for the validity check remains low.

### 2015-01-20 PLDI15_Response/response.txt

We thank the reviewers for their detailed reviews and comments.

Complexity of contracts  (Reviewers ...)
-----------------------

We believe the difficulty of writing contracts is substantially simpler than
choosing the correct store consistency level since developers do not need to
understand the details of the underlying data store.

Moreover, while it is necessary to translate integrity constraints in the
application to consistency constraints as contracts, the tremendous
advantage that accrues is that these contracts are expressed over an
abstract system model, resulting in generic RDT specifications that can be
realized on different data stores (as our implementation illustrates) with
varied consistency lattices (as pointed out by ReviewerE).

The Quelea contract language itself is built upon *only* three primitive,
well-understood relations -- vis, so and sameobj, along with standard
propositional logical and relational operators.  In the presence of
coordination-free transactions with weaker isolation levels, the choice of
correct consistency level is non-trivial. Quelea incorporates transactions
with a single additional primitive relation (sametxn).

< provide an example using syntactic sugar >

Comparison to previous work (Reviewers ...)
---------------------------

Our work differs significantly from Bailis2013b and Burckhardt2014.

Burckhardt2014 describes a formal framework for expressing, verifying and
comparing different forms of eventual consistency, and is not meant for
practical programming over ECDS like Quelea. Moreover, unlike
Burckhardt2014, Quelea focuses on *automatic* verification of RDT
specifications, which is essential for practicality, and additionally
incorporates support for programming with coordination-free transactions.

Bailis2013b describes strategies to realize stronger consistency levels on
top of an eventually consistent store, but does not consider
parameterization of application consistency constraints over arbitrary data
store implementations.  In contrast, Quelea provides a language framework to
describe application-level consistency requirements and automatically map
them to *some* data store implementation. As such, Quelea's contract
classification is independent of how various consistency levels are actually
implemented in the data store.  Moreover, Bailis2013b does not incorporate
transactions, which Quelea does.

Compilation Time  (Reviewers ...)
----------------

Our contract classification through the solver took on average took 11.5 ms
across the benchmarks (this statistic is also available on p.11 of the
paper).  We expect this SMT solver overhead to remain low as the application
scales. This is because we do not analyze the RDT/transaction implementation
to obtain the verification conditions but only analyze user-provided
contracts.  Moreover, our contract language artifacts (e.g: primitive
relations, effects), are encoded as uninterpreted symbols in SMT without the
need for expensive theories, such as linear arithmetic.  Hence, the state
space for the validity check remains low.

The time to classify a contract is proportional to the number of consistency
levels in the partially ordered lattice. For example, Doug Terry et al.'s
Session Guarantees for Weakly Consistent Replicated Data (cite ...)
introduced 4 incomparable session guarantees, and the consistency level for
an operation is any combination of the 4 session guarantees. The
classification here, at worst, requires 4! = 24 comparisons (SMT validity
tests). Since each comparison is very cheap, the solution remains scalable.

Review A
--------

The challenge with a dynamic strategy for enforcing contracts is the
translation of contracts, which axiomatically describe correct execution to
a validity check over the dynamic execution state. Such a translation is
non-obvious.

@@

We used the same number of shim layer nodes for the results in Figure 9
a,b,c -- 5. Shim Layer nodes are soft-state entities and can be elastically
scaled on demand. All the benchmarks except Twissandra and RUBiS were
written from scratch.

Review B
--------

Bounded staleness could be accommodated into the contract language, but
contract classification becomes undecidable since the theory of integer
arithmetic is undecidable.

The RDT specification language does to some extent guard against
malicious/incorrect code. In particular, RDT operations are side-effect
free.  While it is possible that the reduction can end up not terminating,
it cannot corrupt system state.

@@

A crashed operation does not affect the durable log. An operation is added to
the log only if it is successful. An operation is considered successful if it
produces a result.

R+ could be a total relation, but R+ is still useful because it is not
required to be a total relation. For example, consider a causal
consistency (CC) contract defined as:

@@

 hbo \subseteq vis

where hbo = (soo U vis)+. Because of our lenient encoding of “+", hbo can be
any superset of the transitive closure of (soo U vis). Nevertheless, all we
can assume about hbo is that it contains the  transitive closure of (soo U
vis); there is no proof of membership for other elements in hbo. 
Consequently, by asserting hbo \subseteq vis, the strongest guarantee that
causal consistency  contract is offering is that the transitive closure of
(soo U vis) is contained in vis. An application  contract (\psi) which
requires more than the transitive closure of (soo U vis) to be contained in
vis  cannot be satisfied under CC because the following implication is
simply not valid even under our lenient definition of “+“ :

@@

 \Delta |- (hbo \subseteq vis) => \psi

\Delta, as usual, denotes the conjunction of basic axioms about primitive
relations.

Review E
--------

Our idea is to parameterize the classification procedure for each target
data store. Hence, if multiple partially ordered levels are appropriate, the
classification procedure can be encoded to choose the desired consistency
level based on external knowledge such as price.

### 2015-01-20 PLDI15_Response/response.txt

We thank the reviewers for their detailed reviews and comments.

Complexity of contracts  (Reviewers ...)
-----------------------

We believe the difficulty of writing contracts is substantially simpler than
choosing a correct store consistency level, since developers do not need to
understand the details of the underlying data store to define consistency
properties relevant for the application.

Moreover, while it is necessary to translate application-specific integrity
constraints to consistency constraints in the form of contracts, the
tremendous advantage that accrues is that these contracts are expressed over
an abstract system model, resulting in generic RDT specifications that can
be realized on different data stores (as our implementation illustrates)
with varied consistency lattices (as pointed out by ReviewerE).

The Quelea contract language itself is built upon *only* three primitive,
well-understood relations -- vis, so and sameobj, along with standard
propositional logical and relational operators.  In the presence of
coordination-free transactions with weaker isolation levels, the choice of
choosing the correct consistency level is non-trivial. Note also that Quelea
incorporates transactions with a single additional primitive relation
(sametxn).

### 2015-01-21 PLDI15_Response/response.txt

We thank the reviewers for their detailed reviews and comments.

Complexity of contracts  (Reviews B,C,D,E)
-----------------------

@@

We believe the difficulty of writing contracts is substantially simpler than
choosing a correct store consistency level, since developers do not need to
understand the details of the underlying data store to define consistency
properties relevant for the application.

Moreover, while it is necessary to translate application-specific integrity
constraints to consistency constraints in the form of contracts, the tremendous
advantage that accrues is that these contracts are expressed over an abstract
system model, resulting in generic RDT specifications that can be realized on
different data stores (as our implementation illustrates) with varied
consistency lattices (as pointed out by ReviewerE).

The Quelea contract language itself is built upon *only* three primitive,
well-understood relations -- vis, so and sameobj, along with standard
propositional logical and relational operators.  In the presence of
coordination-free transactions with weaker isolation levels, the choice of
choosing the correct consistency level is non-trivial. Note also that Quelea
incorporates transactions with a single additional primitive relation
(sametxn).

Comparison to previous work (Reviews C,D)
---------------------------

@@

Bailis2013b describes strategies to realize stronger consistency levels on top
of an eventually consistent store, but does not consider parameterization of
application consistency constraints over arbitrary data store implementations.
In contrast, Quelea provides a language framework to describe application-level
consistency requirements and automatically map them to *some* data store
implementation. As such, Quelea's contract classification is independent of how
various consistency levels are actually implemented in the data store.
Moreover, Bailis2013b does not incorporate transactions, which Quelea does.

Compilation Time  (Reviews A,D)
----------------

Our contract classification through the solver took on average took 11.5 ms
across the benchmarks (this statistic is also available on p.11 of the paper).
We expect this SMT solver overhead to remain low as the application scales.
This is because we do not analyze the RDT/transaction implementation to obtain
the verification conditions but only analyze user-provided contracts.
Moreover, our contract language artifacts (e.g: primitive relations, effects),
are encoded as uninterpreted symbols in SMT without the need for expensive
theories, such as linear arithmetic.  Hence, the state space for the validity
check remains low.

The time to classify a contract is proportional to the number of consistency
levels in the partially ordered lattice. For example, [[Doug Terry et al.'s
Session Guarantees for Weakly Consistent Replicated Data, PDIS '94]] introduced
4 incomparable session guarantees, and the consistency level for an operation
is any combination of the 4 session guarantees. The classification here, at
worst, requires 4! = 24 comparisons (SMT validity tests). Since each comparison
is very cheap, the solution remains scalable.

@@

The challenge with a dynamic strategy for enforcing contracts is the
translation of contracts, which axiomatically describe correct execution to
a validity check over the dynamic execution state. Such a translation is
non-obvious.

We used the same number of shim layer nodes for the results in Figure 9
a,b,c -- 5. Shim Layer nodes are soft-state entities and can be elastically
scaled on demand. All the benchmarks except Twissandra and RUBiS were
written from scratch.

Review B
--------

Bounded staleness, which depends on wall clock time, is not incorporated in the
contract language; as integer arithmetic, in general, is undecidable, its
inclusion would make the classification procedure undecidable.

The standard benchmarks such as TPC-C are designed for traditional relational
database systems whose specifications dictate relational tables, joins and
transactions with strong isolation requirements, which are not readily
supported by NoSQL stores. As such these traditional benchmarks such as TPC are
not used to analyze and compare ECDS performance. On the other hand, since
Quelea was realized on top of Cassandra, we were able to obtain a point of
comparison with respect to vanilla Cassnadra (P.10, Column.2, para.3).

R+ could be a total relation, but R+ is still useful because it is not
*required* to be a total relation. For example, consider a causal
consistency (CC) contract defined as hbo \subseteq vis where hbo = (soo U
vis)+. Though our definition hbo can be a superset of transitive closure (TC)
of (soo U vis), we do not have a *proof* of inclusion of elements other than
the ones in the TC. As such an application contract (\psi) which requires more
than CC (say total order) would not be satisfied since the implication \Delta
|- (hbo \subseteq vis) => \psi is not valid.

### 2015-02-18 PLDI15_Response/response.txt

The challenge with a dynamic strategy for enforcing contracts is the
translation of contracts, which axiomatically describe correct execution to
a validity check over the dynamic execution state. Such a translation is
non-obvious.

We used the same number of shim layer nodes for the results in Figure 9
a,b,c -- 5. Shim Layer nodes are soft-state entities and can be elastically
scaled on demand. All the benchmarks except Twissandra and RUBiS were
written from scratch.

Review B
--------

Bounded staleness, which depends on wall clock time, is not incorporated in the
contract language as integer arithmetic, in general, is undecidable; its
inclusion would make the classification procedure undecidable.

The standard benchmarks such as TPC-C are designed for traditional
relational database systems whose specifications dictate relational tables,
joins and transactions with strong isolation requirements, which are not
readily supported by NoSQL stores. As such these kinds of benchmarks are not
used to analyze and compare ECDS performance. On the other hand, since
Quelea was realized on top of Cassandra, we were able to obtain a point of
comparison with respect to vanilla Cassnadra (P.10, Column.2, para.3).

### 2015-03-17 motivation.tex
Constructing getBal contract in stages.

\noindent For \cf{getBalance}, we construct the following contract ($\cv_{gb}$):

@@

\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$
is an effect that was emitted either by a \cf{deposit} or a \cf{withdraw}
operation. If a \cf{withdraw} $b$ is visible to \cf{getBalance} $\cureff$, then
all \cf{deposit} operations $a$ visible to $b$ should also be visible to
$\cureff$. Intuitively, this prevents negative balance anomalies. Our contract
language provides operators to compose relations. The syntax $(R_1 \cap
R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
above contract says that if a \cf{deposit} or a \cf{withdraw} operation
precedes a \cf{getBalance} operation in session order, and is applied on the
same object as the \cf{getBalance} operation, then it must be the case that the
\cf{getBalance} operation witnesses the effects of the preceding operations.

>>>>

\noindent To prevent \cf{getBalance} from ever showing negative balance, it is
necessary to prevent the scenario depicted in
fig.~\ref{fig:negativeBalanceAnomaly}. Let $\cureff$ stand for the effect
emitted by the \cf{getBalance} operation. If the effect ($b$) of a withdraw
operation is visible to $\cureff$, and the effect ($a$) of a deposit operation
is visible to the effect ($b$) of the withdraw operation, then it must be the
case that $a$ is also visible to $\cureff$. A contract ($\cv_{gb}^1$) for
\cf{getBalance} operation precisely captures this application-level consistency
requirement:

\vspace{-1em}
\begin{smathpar}
\begin{array}{l}
\forall (a:\rcf{deposit}), (b:\rcf{withdraw}). \\
\qquad\qquad (\vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff})
\end{array}
\end{smathpar}

\noindent To prevent the missing update anomaly described in
fig.~\ref{fig:missingUpdateAnomaly}, it is necessary for a \cf{getBalance}
operation on a bank account to witness the effects of all previous \cf{deposit}
and \cf{withdraw} operations performed on the same bank account in the same
session. We can conceive an alternative contract ($\cv_{gb}^2$) for
\cf{getBalance} to express this consistency requirement:

\vspace{-1em}
\begin{smathpar}
\begin{array}{l}
\forall (c: \rcf{deposit} \vee \rcf{withdraw}). \\
\qquad \qquad((\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff})
\end{array}
\end{smathpar}

\noindent Our contract language provides operators to compose relations. The
syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The
above contract ($\cv_{gb}^2$) says that if a \cf{deposit} or a \cf{withdraw}
operation precedes a \cf{getBalance} operation in session order, and is applied
on the same object as the \cf{getBalance} operation, then it must be the case
that the \cf{getBalance} operation witnesses the effects of the preceding
operations.

\noindent To final contract ($\cv_{gb}$) of the \cf{getBalance} operation is
merely a conjunction of the previous two versions ($\cv_{gb}^1$ and
$\cv_{gb}^2$):

@@

\noindent Intuitively, this prevents both the \cf{getBalance} anomalies
described in figs.~\ref{fig:negativeBalanceAnomaly} and
\ref{fig:missingUpdateAnomaly} from ever occurring.
